Lyman-$\alpha$ photons through rotating outflows

### 2018-08-10 paper/paper.tex
edit

@@

We use (CLARA) \citep{CLARA}, a Monte Carlo code that follows the
propagation of individual photons through a neutral Hydrogen medium
characterized by  its temperature, velocity field and global optical depth.
The code assumes an homogeneous density throughout the simulated volume.
In the current implementation we neglect the influence of dust.
Our basic model is an spherical distribution of neutral hydrogen,
an approximation commonly used in the literature, as it explains a
wide variety of observational features
\citep{Ahn03,Verhamme06,Dijkstra06}.

The velocity field we use captures both outflows and rotation.
Outflows are described by a Hubble-like radial velocity profile with
the velocity magnitude increasing linearly with the radial
coordinate; this model is fully characterized by $v_{\rm out}$, the
radial velocity at the sphere's surface.
Rotation follows a solid body rotation profile, which is fully
characterized by $v_{\rm rot}$, the linear velocity at the sphere's surface.

The total velocity field corresponds to the superposition of rotation and
outflows.
The cartesian components take the following form:

@@

For each model setup we follow $10^5$ individual photons generated from
the center of the sphere at the \lya line's center as they propagate
through the volume and finally escape.
We store the frequency and propagation direction for each photon
at its last scattering.

As input parameters we use $\tauh=\{10^5, 10^6, 10^7\}$,
$\vout=\{5,25,50\}\;\kms$ and $\vrot=\{0,50,100\}\;\kms$, for a total of
$27$ models with all the possible parameter combinations.
The values for the outflow velocity are lower than other values used in the
literature to allow for an interplay between the two kinematic features.
Nevertheless, this range of parameters reproduces the main standard
deviation and skewness  of high redshift LAES in recent observations
(\label{subsec:kinematic}).

@@

\cite{Garavito14} presented an analytical model that
accounts for the effects of pure rotation on the
\lya line morphology. 
The basic assumption of their analytical model is that each
differential surface element on the sphere Doppler shifts the photons
that it emits.
In this paper we introduce this ansatz by post-processing the results
of the outflows simulations without rotation.
The frequency of each photon is Doppler shifted as follows

\begin{equation}
x' = x + \frac{\vec{v}_{\rm rot}\cdot \hatk}{\vth}
\label{eq:shift_x}
\end{equation}
%
where $x'$ is the photon's new adimensional frequency, $x$ is the photon's
frequency after being processed only by the outflow, \vrot is the 
rotational velocity at the point of escape of the photon, \hatk is
the photon's direction of propagation and \vth is the thermal
velocity of the sphere. 
In this study we use \vth $= 12.86$ \kms, which corresponds to a
temperature of $T=10^4$ K.   
We fix \vth throughout the paper, its influence is only
of a multiplying constant that can be included without running new
Monte-Carlo simulations, as the code works with the adimensional
frequency $x$. 
However, the thermal velocity is considered a free parameter 
when comparing simulation results against observations.

The approach presented above allows us to produce new \lya spectra and
compare them with the full radiative transfer solution including both
outflows and rotation.

>>>>

Throughout the paper we use a thermal velocity for a neutral Hydrogen
gas of \vth $= 12.86$ \kms, which corresponds to a temperature of
$T=10^4$ K.

@@

We use CLARA \citep{CLARA}, a Monte Carlo code that follows the
propagation of individual photons through a neutral Hydrogen medium
characterized by its temperature, velocity field and global optical depth.
The code assumes an homogeneous density throughout the simulated volume.
In the current implementation we neglect the influence of dust.
Our basic model is an spherical distribution of neutral hydrogen,
an approximation commonly used in the literature that explains a
wide variety of observational features
\citep{Ahn03,Verhamme06,Dijkstra06}.

We use a velocity field that captures both outflows and rotation.
Outflows are described by a Hubble-like radial velocity profile with
the speed increasing linearly with the radial coordinate; this model
is fully characterized by $v_{\rm out}$, the radial velocity at the
sphere's surface. 
Rotation follows a solid body rotation profile fully characterized by
$v_{\rm rot}$ the linear velocity at the sphere's surface.

The total velocity field is the superposition of rotation and
outflows.
The cartesian components take the following form:

@@

For each run we follow $10^5$ individual photons generated from the
origin at the \lya line's center as they propagate
through the volume and finally escape.
We store the frequency and propagation direction for each photon
at its last scattering.

As input parameters we use $\tauh=\{10^5, 10^6, 10^7\}$,
$\vout=\{5,25,50\}\;\kms$ and $\vrot=\{0,50,100\}\;\kms$, for a total of
$27$ models with all the possible parameter combinations.
The values for the outflow velocity are lower than values commonly used in the
literature to allow for an interplay between the two kinematic features.
Nevertheless, this range of parameters reproduces the main standard
deviation and skewness  of high redshift LAES in recent observations
(\label{subsec:kinematic}).

@@

\cite{Garavito14} presented an analytical model that
accounts for the effects of pure rotation on the
\lya line morphology. 
The basic assumption of the model is that each differential surface
element on the sphere Doppler shifts the photons that it emits.
In this paper we introduce this ansatz by post-processing the results
of the outflows simulations without rotation.
The frequency of each photon is Doppler shifted as follows

\begin{equation}
x' = x + \frac{\vec{v}_{\rm rot}\cdot \hatk}{\vth}
\label{eq:shift_x}
\end{equation}
%
where $x'$ is the photon's new adimensional frequency, $x$ is the photon's
frequency after being processed only by the outflow, \vrot is the 
rotational velocity at the point of escape of the photon, \hatk is
the photon's direction of propagation and \vth is the thermal
velocity of the sphere. 
We fix \vth throughout the paper. This factor is a multiplying
constant that could be changed without running new Monte-Carlo
simulations, as the code works with the adimensional frequency $x$.

This semi-analytic model allows us to produce new \lya spectra from
the outflow-only results and compare them with the full radiative
transfer solution including both outflows and rotation.
